Robust Bayesian Analysis of Causal Inference Problems

### 2024-03-09 replies_reviews.txt

----------------------- REVIEW 1 ---------------------

This paper introduces a robustification of a Bayesian approach to indentify 
whether a treatment has an influence on an outcome variable, including possible 
effects of external variables. The model is assumed to be linear (both between 
outcomes and treatments/variables, and between treatment and variables).

Overall, the paper is well written and I believe rigourous (my knowledge of this
particular approach is limited). I would say the technical extension is rather 
simple (considering sets of parameters and performing sampling experiments wihtin 
these sets, followed by a classical Bayesian analysis), and the experiments 
synthetic, hence the most interesting part of the paper appears to me as the one 
trying to identify why going imprecise/robust is good. However the analysis of 
the experiments is a bit cryptic to me. Still, the paper seems ok enough to me to 
be presented and discussed at ECSQARU, and it fits the scope of the conference.

-> Thank you very much for your positive feedbacks. We have modified the 
-> simulation studies to explain the methods in a better way.

Given my limited knowledge of the formal model, most of my relevant comments 
will concern the experiments:

* I do not really understand what is reported in Table 2: what is the quantity 
we read, what should be his ideal value (lower=better, higher=better, fitting a
particular estimate?). Form my understanding, this is the estimate of \beta_T 
once the Bayesian analysis has been achieved. This should be clearly stated. 
The results however seem to show that the Bayesian analysis is doing really fine 
in both cases, and it would also be interesting to report some credibility 
intervals for those. It would also probably be interesting to look what happens 
in case of model/prior misspescification, and how the two approaches would react 
in this case.

-> We propose a robust Bayesian analysis to show its benefits in situation
-> where we have limited data to perform our analysis. We now have modified
-> the text and the table to make it more interpretable.

* Table 3 is a bit frustrating, as 1. we also appear to be in a rebgime where 
classical approaches mostly do fine (maybe finding a regime where imprecision 
works better would be useful?), 2. it is not clear what are the origins/main
contributors to the loss when it is non-null (mostly false positives/false 
negatives or abstention?) and 3. what would give the classical Bayesian approaches 
if one used the proposed loss functions to take decisions, using a rather classical
appraoch of seeing abstention as a peculiar decision in competition with the 
others... albeit by choosing \ell_3=0.5 which is similar to discounted accuracy 
it is quite likely that such an otpion would never be considered by precise methods.

-> Thank you very much for this interesting feedback. As we mentioned earlier,
-> we are particularly focusing on the situations where we have limited data.
-> Since it is a regressional framework, the analyses tends to support
-> classical approaches for $n \ge 50$. We have modified the simulation
-> setting to show results for $25 \le n \le 75 $, to reflect a regime where
-> imprecision is more useful.

Minor comments:

* Equations (2) and (3): I am not sure what to do subject-specific random variable? 
How can we interpret a single individual experiment as a random quantity (I know 
this is classical in statistics and that "in theory" everything goes fine, but in
practice this seems weird).

-> The random variables for subject specific
-> quantities, such as Yi and Ti, denote quantities before they are observed.
-> Since these quantities, before observation, are uncertain, they must be treated
-> as random variables. For instance, the decision to treat or not, Ti, could be
-> based on some bio markers, which are inherently random due to noise.
-> The uncertainty about the response Yi could be more epistemic in nature,
-> although there could also be random factors that influence the response.
-> The resulting subject specific values of delta_i correspond then to a
-> mathematical formulation of treatment effect heterogeneity.

* I am not sure Table 1 information should be a table rather than in-line texts. 
It seems to me like its taking space without adding much structure/info.

-> We now have modified the table accordingly.

----------------------- REVIEW 2 ---------------------

The authors focus on causal inference, and the goal here is to enlarge the usual
Bayesian analysis so that a set of probability distributions is contemplated and 
so that robustness can be assessed. This is a relevant issue (even though it is
presented as if its relevant is obvious, and some words connecting with the bursting
literature on causal thinking in AI would be really valuable). And the paper is 
well written and easy to follow. The contribution is somewhat narrow as it basically
consists in taking an existing model and adding a set of probabilities to it, so 
as to examine robustness and sensitivity. This makes sense but the idea seems quite
natural and does not seem to bring technical insight (perhaps this should be 
discussed: is there any technical issue that is challenging in doing the extension
proposed in the paper?). The weak part is the validation, as it is based on apparently
simple simulations. It would be a lot more convincing to use real data; given the
amount of data available today, it would be strange not to find some dataset that 
is appropriate. Any thoughts?

-> Thank you for the comment. We agree it would be better to use real data.
-> Unfortunately we cannot fully implement this suggestion for the conference paper,
-> but we have acknowledged this issue in the conclusion, and we shall definitely take
-> your suggestion forward in future work.

Concerning the text, a few suggestions:
- The abstract is rather long and wordy; perhaps it could be shortened and made sharper.

-> We agree. We have made the abstract a bit shorter.

- A few places: "can not" instead of "cannot".
- Expressions (20) and (21) should have ending punctuation.
- Page 9: "under estimates" instead of "underestimates"; also, expression "the near
vacuous of set of" seems incorrect.
- References:  "Bayesian" instead of "bayesian".

-> Thank you very much for pointing out the typos. We have corrected those accordingly.

----------------------- REVIEW 3 ---------------------

This is paper focuses on Bayesian hierarchical methods for the estimation of causal effects. The authors adopt a standard regression model for causal estimation and a Bayesian hierarchical model based on that model. The goal is to cope with data scarcity, and a robust approach consisting of a sensitivity analysis wrt the priors is used. This gives a procedure for covariate selection.  A utility approach is also sketched to evaluate whether the available data are sufficient for a robust decision. The method is tested on a synthetic model, and the results compete with existing approaches.

I understand the work of the authors as an application of ideas from robust Bayes and imprecise probability theories to causal analysis. This looks very promising to me. Even if I regard the present work as a first attempt in this direction, the paper might trigger interesting discussions at the conference.

-> Thank you for your positive comments.

I miss some discussions about the optimisation method. I understand everything here is based on sampling, but the differences between standard approaches should also be emphasised. I also don't understand whether the bound approximation can be characterised (e.g., inner approx).

-> Thank you for the comment. We agree it would be better to investigate further theoretical
-> properties and obtain approximations of the posterior estimates.
-> Unfortunately we cannot finish the implementation for the conference paper,
-> but we have acknowledged this issue in the conclusion, and we shall definitely take
-> your suggestion forward in future work.

More generally speaking, the authors might also be interested in the works about the bounding of counterfactuals. Sampling methods have been proposed for that (see various recent works by Bareinboim and colleagues and also Zaffalon and colleagues).

-> Thank you for this interesting insight, currently this is beyond the scope of our
-> contribution but we will definitely take your suggestion for the extension of the
-> paper.

### 2024-03-21 JASA-template.tex
Move Jochen to acknowledgements.

\usepackage[numbers]{natbib}
\usepackage{url} % not crucial - just used below for the URL 
\usepackage{doi}

@@

\title{Robust Bayesian Analysis of Causal Inference Problems}
\author{Tathagata Basu$^1$}
\email{tathagatabasumaths@gmail.com}
\author{Matthias C.~M.~Troffaes$^2$}
\email{matthias.troffaes@durham.ac.uk}
\author{Jochen Einbeck$^{2,3}$}
\email{jochen.einbeck@durham.ac.uk}
\address{$^1$Civil and Environmental Engineering, University of Strathclyde}
\address{$^2$Department of Mathematical Sciences, Durham University}
\address{$^3$Durham Research Methods Centre}

\begin{abstract}
Causal inference using observational data is an important aspect in
many fields such as epidemiology, social science, economics, etc. In
particular, our goal is to find the treatment effect on the subjects
along with the causal links between the variables and the outcome.
However, estimation for such problems are extremely 
difficult as the treatment effects may vary from subject 
to subject and modelling the underlying heterogeneity explicitly makes the 
problem practically unsolvable. Another issue we often face is the 
dimensionality of the problem and we need to find a subset of 
explanatory variables to initiate the treatment. However, currently variable selection methods
tend to maximise the predictive performance of the outcome model only. This can be problematic in the case of limited information. 
As the consequence of mistreatment can be harmful. 
So, in this paper, we suggest a general framework with
robust Bayesian analysis which accounts for abstention in deciding
an explanatory variable in the high dimensional regression model. 
To achieve that, we consider a set of spike and slab priors 
through prior elicitation to obtain robust estimates for
both the treatment and outcome model. We are specifically interested 
in the sensitivity of the treatment effect in the high dimensional causal inference
as well as the identifying the confounder variables by means of variable selection. However, indicator
based confounder selection can be deceptive in some cases. Especially, 
when the predictor is strongly associated with either the treatment or 
the outcome. This increases the posterior expectation of the selection
indicators. To avoid that we apply a post-hoc selection scheme
which successfully remove negligible non-zero effects from the model
attaining a smaller set of confounders. Finally, we illustrate
our result using synthetic \alert{and real} dataset.

@@

\maketitle

\section{Introduction}\label{sec:intro}

In causal inference, we are interested in estimating the causal
effect of independent variables on a dependent variable. Ideally,
randomised trials are the most efficient way to perform this task.
However, this is not very practical for several reasons; ethical 
concerns, design cost, population size, to name a few. This
leaves us with observational studies which are usually obtained
by means collecting data though surveys or record keeping. But this
can be problematic in the presence of confounders. That is when
the variables are associated with both the treatment and the outcome.
In such cases, we need to be extra cautious as otherwise it will
lead to unwanted bias in the treatment effect estimator \cite{rosenbaum83}.
Several works have been done in order to tackle the presence of
confounder variables. One such work in the topic was 
by \citet{Robins1986ANA} where the author used  a graphical
approach for the identification of the causal parameters.
\citet{rosenbaum1985} suggested the use of a link model to estimate
the propensity scores for all individuals. Later on several other
methods have been proposed based on propensity score matching.
A brief review on such methods can be found in \cite{winship99, stuart10}.

Bayesian approach in causal effect estimation has become a popular
topic in recent days. However, it has been a topic of interest for about 50 
years and one of the earlier works on this can be
found in \cite{rubin1978}. Lately with the notion of high dimensional problems, Bayesian methodologies have become
more appealing. \citet{Crainiceanu2008} proposed a bi-level 
Bayesian model averaging based method for estimating the causal 
effect. \citet{wang2015} suggested BAC (or, Bayesian adjustment for
confounding) where they use an informative prior obtained from
the treatment model and apply them on the outcome model for
estimating causal effect. Several other methods were also
proposed to tackle confounders from the point of view of Bayesian
variable selection such as: \citet{Zigler2014}, \citet{Hahn2018} etc.

In this paper we take inspiration from the approach of \citet{koch2020}, where
they proposed a bi-level spike and slab prior for causal effect 
estimation. They considered a data-driven adaptive approach to
propose their prior which reduce the variance of the causal estimate. 
In our approach, we are interested in the decision theoretic notion of 
variable selection and its effect in causal estimation. We suggest a 
sensitivity analysis based approach where instead of using a single prior, 
we consider a set of priors \cite{BERGER1990303}. This is particularly 
interesting as in many cases, causal effect estimation can be performed 
through a meta analysis and hence robust Bayesian analysis can be beneficial
\cite{raices_cruz22} under severe uncertainty. Moreover, for some problems 
we have to rely on very limited data to perform our Bayesian analysis and 
choosing a data-driven prior can lead to overfitting. Instead we can 
benefit from expert opinion and elicit prior based on empirical evidence.
This also allows us to construct the problem of confounder selection 
in a framework where abstention has a positive gain. To propose our 
framework, we consider a set of continuous spike and slab priors 
\cite{ishwaran2005} for confounder selection and construct a Bayesian 
group LASSO \cite{xu2015} type problem. To perform the prior sensitivity 
analyses, we consider a set of beta priors on the covariate selection 
probability of the spike and slab priors. We use the posteriors of these
covariate selection probability for identifying the confounders. Finally, 
we consider a post-hoc coefficient adjustment method \citet{hahn2015}
to recover sparse estimates. This post hoc treatment also helps us to 
reduce the bias of treatment effect estimation.

The rest of the paper is organised as follows. In \cref{sec:causal}
we give a formal description of causal estimation problem in the
context of linear regression. \cref{sec:bayes} is focused on the
Bayesian analysis of causal inference problems, followed by the
motivation of a robust Bayesian analysis along with our proposed decision 
theoretic framework for confounder (variable) selection. In \cref{sec:sim}, 
we provide our result of simulation studies under different scenarios 
and data analysis with \alert{real data} in \cref{sec:data:analysis}. Finally 
we discuss our findings and conclude this paper in \cref{sec:conc}.

@@

Let an observational study give us the outcomes $Y=(Y_1$, \dots, $Y_n)$ along with 
corresponding treatment indicators $T=(T_1$, \dots, $T_n)$. Then the
treatment 
effect in the population is given by the expectation of the difference
in outcome between the treatment and controls. 
\begin{align}
\delta = \mathbb{E}(Y\mid T =1) - E(Y\mid T=0).
\end{align}
Similarly, individual causal
effect of the treatment $T_i$ on outcome $Y_i$ is given by:
\begin{align}
\delta_i \coloneqq (Y_i\mid T_i=1) - (Y_i\mid T_i=0).
\end{align}
That is, the difference between the outcome when $i$-th subject receives
the treatment and when $i$-th subject remain as a control.

In theory, both of these quantities exist.
However, we can not observe $(Y_i\mid T_i=1)$ and $(Y_i\mid T_i=0)$
simultaneously for the $i$-th individual. Instead, we can estimate
average causal effect of the treatment $T$ by calculating the averaged
outcome of all the subjects those received the treatment and
all the subjects those remained as control.
\begin{align}
\hat{\delta} \coloneqq 
\frac{\sum_{i=1}^n Y_i\cdot\mathbb{I}(T_i=1) - 
	\sum_{i=1}^n Y_i\cdot\mathbb{I}(T_i=0)}{n}.
\end{align}
However, this relies on an important assumption that the treatment effect
on the $i$-th subject given that they received the treatment is
equal to the treatment effect when they remain as the control
\cite{winship99}.

\subsection{Regression Model}
Regression methods are widely used in causal effect estimation. The
main idea behind these regression methods is to remove the
correlation between the treatment indicator and the error term
\cite{winship99,HECKMAN1985}. To do so, we rely on $p$ different observed quantities
or predictors denoted by $X\coloneqq$ $[X_1$, \dots, $X_n]^T$. Now, let
$\beta \coloneqq (\beta_1$, \dots, $\beta_p)$ denotes the vector of regression
coefficients. Then we can define a linear model for the outcome
so that
\begin{equation}
    Y_i = \beta_{T} T_i + \beta_0 + X_i\beta + \epsilon_i
\end{equation}
where $\epsilon_i\sim \mathcal{N}(0, \sigma^2)$. Clearly, when
the underlying true outcome model is linear,
\begin{equation}
    \delta = \beta_{T}\quad\text{and hence } 
    \hat{\delta} - \delta = \hat{\beta}_T - \beta_T.
\end{equation}
\begin{figure}
	\centering
	\begin{tikzpicture}[params/.style={circle, draw=green!60, fill=green!5, very thick, minimum size=7mm}]
	\node (1) at (0,0) {Treatment ($T$)};
	\node (2) [right = of 1] {Outcome ($Y$)};
	\node (4) [below = of 1] {Predictors ($X$)};

	\path (1) edge[->]  (2);
	\path (1) edge[<-]  (4);
	\path (2) edge[<-] (4);
	\end{tikzpicture}
	\caption{Confounding in causal models.}
	\label{fig:confounding}
\end{figure}

In the presence of confounders we also need to consider the
association between the treatment indicators and the predictors.
\citet{koch2020} suggested the use of a probit link function to 
construct the regression model. This way, we can
specify the conditional probability that subject $i$ receives the treatment through a linear model. 
That is, for another vector of regression coefficients 
$\gamma\coloneqq(\gamma_1, \cdots, \gamma_p)$ we define
\begin{align}
    P(T_i=1\mid X_i) \coloneqq \Phi(\gamma_0+X_i\gamma)
\end{align}
where $\Phi(\cdot)$ denotes the cumulative distribution function
of a standard normal distribution.

This allows us to define intermediate latent variables as suggested
by \citet{albert93}
\begin{equation}
	T_i^* = \gamma_0 + X_i\gamma +u_i
\end{equation}
where, $u_i\sim\mathcal{N}(0,1)$. Therefore, $T_i=1$ if $T_i^*>0$ and
$T_i=0$ if $T_i^*\le0$.

Now, following the approach of \citet{koch2018}, we define an adjusted
output vector $W\coloneqq(Y, T^*)^T$ and corresponding $2n\times(2p+3)$ dimensional design matrix
\begin{align}
    Z &=
    \begin{bmatrix}
     X_O & 0 \\
     0 & X_T
    \end{bmatrix},
\end{align}
where, $X_O = [T, 1_n, X]$ and $X_T = [1_n, X]$. Then, assuming a
Gaussian error term, we have the following likelihood distribution
\begin{align}
W\mid Z, \alpha, \beta, \gamma, \sigma^2 \sim\normal\left(Z\nu, \Sigma\right)\label{eq:like:group},
\end{align}
where $\nu = (\beta_T, \beta_0, \beta, \gamma_0, \gamma)^T$ and
\begin{align}
\Sigma &=
\begin{bmatrix}
\sigma^2{I}_n & 0 \\
0 & {I}_n
\end{bmatrix}.
\end{align}

@@

The likelihood formation given by \cref{eq:like:group} gives us
a foundation for Bayesian group LASSO 
\cite{xu2015} type model and look into the posterior selection
probability associated with the $j$-th predictor. There are several
ways to construct spike and slab priors which achieve 
variable selection. In our case, we consider a continuous type
\cite{ishwaran2005} prior for faster posterior
computation.

@@

Let, $\pi_j$ denote the prior probability that the $j$-th
predictor is not associated with either the outcome or the 
treatment. That is, 
\begin{equation}
	\pi_j = P\left((\beta_j,\gamma_j)=(0,0)\right).
\end{equation}
Then we can define a spike and slab group LASSO so that:
for $1\le j\le p$,
\begin{align}
(\beta_j,\gamma_j)^T \mid \pi_{j}, \sigma^2 &\sim 
(1-\pi_{j})\normal\left( (0,0)^T, 
\tau_1^2\begin{bmatrix}
\sigma^2 & 0 \\
0 & 1
\end{bmatrix}\right)
+ \pi_{j} \normal\left(0, 
\tau_0^2\begin{bmatrix}
\sigma^2 & 0 \\
0 & 1
\end{bmatrix}\right)\\
%\beta_T, \beta_0\mid \sigma^2 &\sim \normal(0, \sigma^2)\\
%\gamma_0 &\sim \normal(0,1)\\
\sigma^2&\sim \text{InvGamma}(a, b)\\
\pi_{j} &\sim\text{Beta}\left(sq_j, s(1-q_j)\right).
\end{align}

We fix sufficiently small $\tau^2_0$
$(1\gg\tau_0^2>0)$ so that  $(\beta_j, \gamma_j) = (0,0)$ has its probability mass 
concentrated around zero. Therefore, this represents the spike component of our prior specification. 
To construct the slab component, we consider $\tau_1^2$ to be large so that $\tau_1^2\ge 1$. This allows the prior for $(\beta_j, \gamma_j)\not=(0,0)$ to be flat. Inverse-Gamma is a natural choice
for the variance of the Gaussian noise because of conjugacy. To
make the prior flat, we consider $1\gg a, b >0$.

As indicated earlier, $\pi_j$ is used as the rejection probability
of the $j$-th predictor and we use a beta prior to specify 
these rejection probabilities where  $q_j$ represents our prior expectation of the rejection probability ($\pi_j$) and `$s$' acts as 
a concentration parameter.
For the intercept terms of the outcome model and the causal effect, 
we consider a sufficiently flat prior so that 
$\beta_0. \beta_T\sim \normal(0,\sigma^2)$. Similarly, for the
intercept term in the treatment model, we consider 
$\gamma_0\sim \normal(0,1)$.

In \cref{fig:regress}, we show a probabilistic graphical representation
of our hierarchical model. In the figure, grey circular nodes represent the
prior hyper-parameters which will be used for sensitivity analysis
of the model. The transparent circular nodes are used to denote
the modelling parameters which are our quantities of interest. 
The observed quantities are denoted with transparent rectangular
nodes. We also use a grey rectangular node to denote the intermediate
latent variable $T^*$. We use directed edges to denote the
relationship between different nodes. However, we use a dashed
edge between $X$ and $T$ as we they are related through the latent
variable $T^*$. This dashed edge also establish the notion of
confounding as shown in \cref{fig:confounding}.

>>>>

\usepackage[numbers]{natbib}
\usepackage{url} % not crucial - just used below for the URL 
\usepackage{doi}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{subcaption}

@@

\title{Robust Bayesian Analysis of Causal Inference Problems}
\author{Tathagata Basu$^1$}
\email{tathagatabasumaths@gmail.com}
\author{Matthias C.~M.~Troffaes$^2$}
\email{matthias.troffaes@durham.ac.uk}
\address{$^1$Civil and Environmental Engineering, University of Strathclyde}
\address{$^2$Department of Mathematical Sciences, Durham University}

\begin{abstract}
Causal inference using observational data is an important aspect in
many fields such as epidemiology, social science, economics, etc. In
particular, our goal is to find the treatment effect on the subjects
along with the causal links between the variables and the outcome.
However, estimation for such problems are extremely 
difficult as the treatment effects may vary from subject 
to subject and modelling the underlying heterogeneity explicitly makes the 
problem practically unsolvable. Another issue we often face is the 
dimensionality of the problem and we need to find a subset of 
explanatory variables to initiate the treatment. However, currently variable selection methods
tend to maximise the predictive performance of the outcome model only. This can be problematic in the case of limited information. 
As the consequence of mistreatment can be harmful. 
So, in this paper, we suggest a general framework with
robust Bayesian analysis which accounts for abstention in deciding
an explanatory variable in the high dimensional regression model. 
To achieve that, we consider a set of spike and slab priors 
through prior elicitation to obtain robust estimates for
both the treatment and outcome model. We are specifically interested 
in the sensitivity of the treatment effect in the high dimensional causal inference
as well as the identifying the confounder variables by means of variable selection. However, indicator
based confounder selection can be deceptive in some cases. Especially, 
when the predictor is strongly associated with either the treatment or 
the outcome. This increases the posterior expectation of the selection
indicators. To avoid that we apply a post-hoc selection scheme
which successfully remove negligible non-zero effects from the model
attaining a smaller set of confounders. Finally, we illustrate
our result using synthetic dataset.

@@

\maketitle
\section{Introduction}\label{sec:intro}

In causal inference, we are interested in estimating the causal
effect of independent variables on a dependent variable. Ideally,
randomised trials are the most efficient way to perform this task.
However, this is not always practical for several reasons; ethical 
concerns, design cost, population size, to name a few. This
leaves us with observational studies which are usually obtained
by means of collecting data though surveys or record keeping. But this
can be problematic in the presence of confounders, which are variables associated with both the treatment and the outcome.
In such cases, we need to be extra cautious as otherwise it will
lead to unwanted bias in the treatment effect estimator \cite{rosenbaum83}.
Several works have been done in order to tackle the presence of
confounder variables. One such work in the topic was 
by Robins \cite{Robins1986ANA} where the author used  a graphical
approach for the identification of the causal parameters.
Rosenbaum and Robin \cite{rosenbaum1985} suggested the use of a link model to estimate
the propensity scores for all individuals. Later on several other
methods have been proposed based on propensity score matching.
A brief review on such methods can be found in \cite{winship99, stuart10}.

The Bayesian approach in causal effect estimation is a popular
strategy in the field and one of the earlier works on this can be
found in \cite{rubin1978}. Lately, with the rise of high dimensional 
data, Bayesian methodologies have become
more appealing. Crainiceanu \text{et al.~}\cite{Crainiceanu2008} proposed a bi-level 
Bayesian model averaging based method for estimating the causal 
effect. Wang \text{et al.~}\cite{wang2015} suggested BAC (or, Bayesian adjustment for
confounding) where they use an informative prior obtained from
the treatment model and apply them on the outcome model for
estimating causal effect. Several other methods were also
proposed to tackle confounders from the point of view of Bayesian
variable selection, see for instance \cite{Zigler2014,Hahn2018} among others.

In this paper we take inspiration from the approach of Koch \text{et al.~}\cite{koch2020}, who proposed a bi-level spike and slab prior for causal effect 
estimation. They considered a data-driven adaptive approach to
propose their prior which reduces the variance of the causal estimate. 
In our approach, we perform a 
sensitivity analysis based approach where instead of using a single prior, 
we consider a set of priors \cite{BERGER1990303}. This is particularly 
interesting as in many cases, causal effect estimation can be performed 
through a meta analysis and hence robust Bayesian analysis \cite{raices_cruz22} 
can be beneficial under severe uncertainty. Moreover, for some problems 
we have to rely on very limited data to perform our Bayesian analysis and 
inference may not be reliable in presence of heteroscedasticity within
the data. Instead, we use expert opinion and elicit a set of priors 
based on empirical evidence. 
This also allows us to construct the problem of confounder identification 
in a framework where abstention has a relatively positive gain i.e.~
when the cost of further tests/data collection is cheaper than
mistreating a subject. To propose our 
framework, we consider a set of continuous spike and slab priors 
\cite{ishwaran2005} for confounder identification and construct a Bayesian 
group LASSO \cite{xu2015} type problem. To perform the prior sensitivity 
analysis, we consider a set of beta priors on the covariate selection 
probability of the spike and slab priors. We use the posteriors of this
covariate selection probability for identifying the confounders. Finally, 
we consider a post-hoc coefficient adjustment method \cite{hahn2015}
to recover sparse estimates associated with either the outcome or the
treatment model.

The rest of the paper is organised as follows. In \cref{sec:causal}
we give a formal description of the causal estimation problem in the
context of linear regression. \Cref{sec:bayes} is focused on the
Bayesian analysis of causal inference problems, followed by the
motivation of a robust Bayesian analysis along with our proposed decision 
theoretic framework for confounder (variable) selection. In \cref{sec:sim}, 
we provide results of simulation studies under different scenarios 
and show the possible applications in real life problems. Finally, 
we discuss our findings and conclude this paper in \cref{sec:conc}.

@@

Let an observational study give us the outcomes $Y=(Y_1, \dots, Y_n)$ along with 
corresponding treatment indicators $T=(T_1, \dots, T_n)$. Then the
treatment 
effect in the population is given by the expectation of the difference
in outcomes between the treatment and controls:
\begin{align}
	\delta = \mathbb{E}(Y\mid T =1) - \mathbb{E}(Y\mid T=0).
\end{align}
Similarly, the individual causal
effect of treatment $T_i$ on outcome $Y_i$ is given by:
\begin{align}
	\delta_i \coloneqq \mathbb{E}(Y_i\mid T_i=1) - \mathbb{E}(Y_i\mid T_i=0).
\end{align}
That is, we are interested in the difference between the outcomes when the $i$-th subject receives
the treatment and when 
%$i$-th subject 
it remains as a control.

In theory, both of these quantities exist.
However, we cannot observe $\mathbb{E}(Y_i\mid T_i=1)$ and $\mathbb{E}(Y_i\mid T_i=0)$
the average causal effect of the treatment $T$ by calculating the averaged
outcome of all the subjects that received the treatment and
all the subjects that remained as control:
\begin{align}
	\hat{\delta} \coloneqq 
	\frac{\sum_{i=1}^n Y_i\cdot\mathbb{I}(T_i=1) - 
		\sum_{i=1}^n Y_i\cdot\mathbb{I}(T_i=0)}{n}.
\end{align}
However, this relies on an important assumption that the treatment effect
on the $i$-th subject given that they received the treatment is the same as the (counterfactual) treatment effect when they remain as control \cite{winship99}.

\subsection{Regression Model}
Regression methods are widely used in causal effect estimation. The
main idea behind these regression methods is to remove the
correlation between the treatment indicator and the error term
\cite{winship99,HECKMAN1985}. To do so, we rely on $p$ different observed quantities
or predictors denoted by $X\coloneqq$ $[X_1^T, \dots, X_n^T]^T$
where each $X_i\in\mathbb{R}^p$.
Each $X_i$ is treated as a $p$-dimensional row vector,
so $X$ is a $n\times p$ matrix.
Now, let
$\beta \coloneqq (\beta_1$, \dots, $\beta_p)^T$ denote the vector of regression
coefficients
related to the predictors, and $\beta_T$ denote a regression coefficient related to the treatment.
Then we can define a linear model for the outcome
so that
\begin{equation}
	Y_i =  T_i \beta_{T} + X_i\beta + \epsilon_i
\end{equation}
where $\epsilon_i\sim \mathcal{N}(0, \sigma^2)$. Clearly, when
the underlying true outcome model is linear with respect to the treatment,
\begin{equation}
	\delta_i = \mathbb{E}(Y_i\mid T_i =1) - \mathbb{E}(Y_i\mid T_i=0) = \beta_{T}.
\end{equation}
In the presence of confounders we also need to consider the
association between the treatment indicators and the predictors.
In literature, authors often suggest a probit link function to 
construct the regression model. This way, we can
specify the conditional probability that subject $i$ receives the treatment through a linear model. 
That is, for another vector of regression coefficients 
$\gamma\coloneqq(\gamma_1, \cdots, \gamma_p)^T$ we define
\begin{align}
	P(T_i=1\mid X_i) = \Phi(X_i\gamma)
\end{align}
where $\Phi(\cdot)$ denotes the cumulative distribution function
of a standard normal distribution. To incorporate this probit
link function, we assume that we can model the $T_i$ through the following \cite{albert93}:
\begin{align}
    T_i^* &= X_i\gamma +u_i \\
    T_i   &= \mathbb{I}(T_i^*>0)
    =
    \begin{cases}
    1 & \text{if }T_i^*>0 \\
    0 & \text{otherwise}
    \end{cases}
\end{align}
where $u_i\sim\mathcal{N}(0,1)$.

Now, to construct the joint likelihood function, we define an extended
output $2n\times 1$ column vector
$W\coloneqq\left(\begin{smallmatrix}Y \\ T^*\end{smallmatrix}\right)$
and corresponding $2n\times(2p+1)$ dimensional design matrix
\begin{align}
	Z &\coloneqq
        \begin{bmatrix}
           T_1 & X_1 & 0 \\
           \vdots & \vdots & 0 \\
           T_n & X_n & 0 \\
           0 & 0 & X_1 \\
           \vdots & \vdots & \vdots \\
           0 & 0 & X_n
        \end{bmatrix}
        =
	\begin{bmatrix}
		X_O & 0 \\
		0 & X_T
	\end{bmatrix}
\end{align}
where, $X_O = [T, X]$ and $X_T = X$. Then, considering the assumption of
Gaussian error terms, we have the following likelihood distribution
\begin{align}
	W\mid Z, \beta_T, \beta, \gamma, \sigma^2 \sim\normal\left(Z\nu, \Sigma\right)\label{eq:like:group},
\end{align}
where $\nu = (\beta_T, \beta^T, \gamma^T)^T$ and
\begin{align}
	\Sigma &=
	\begin{bmatrix}
		\sigma^2{I}_n & 0 \\
		0 & {I}_n
	\end{bmatrix}.
\end{align}

@@

The likelihood given by \cref{eq:like:group} gives us
a foundation for a Bayesian group LASSO 
\cite{xu2015} type model. This way, we can look into the posterior selection
probability associated with the $j$-th predictor. There are several
ways to construct spike and slab priors which achieve 
variable selection. In our case, we consider a continuous type
\cite{ishwaran2005} prior for faster posterior
computation.

@@

Let $\pi_j$ denote the prior probability that the $j$-th
predictor is associated to the outcome or the 
treatment. That is, 
\begin{equation}
	\pi_j = P\left((\beta_j,\gamma_j)\not=(0,0)\right).
\end{equation}
Then we can define the following hierarchical model for spike and 
slab group LASSO so that,
for $1\le j\le p$,
\begin{align}
	(\beta_j,\gamma_j)^T \mid \pi_{j}, \sigma^2 &\sim 
	\pi_{j}\normal\left( \begin{bmatrix}
		0 \\
		0
	\end{bmatrix}, 
	\tau_1^2\begin{bmatrix}
		\sigma^2 & 0 \\
		0 & 1
	\end{bmatrix}\right)
	+ (1-\pi_{j}) \normal\left(\begin{bmatrix}
		0 \\
		0
	\end{bmatrix}, 
	\tau_0^2\begin{bmatrix}
		\sigma^2 & 0 \\
		0 & 1
	\end{bmatrix}\right)\\
	\beta_T\mid \sigma^2 &\sim \normal\left(0, \sigma^2\right)\\
	\frac{1}{\sigma^2}&\sim \text{Gamma}(a, b)\\
	\pi_{j} &\sim\text{Beta}\left(sq, s(1-q)\right).
\end{align}
In the hierarchical model, we fix sufficiently small $\tau_0$
$(1\gg\tau_0>0)$ so that  $(\beta_j, \gamma_j)$ has its probability mass 
concentrated around zero. Therefore, this represents the spike component of our prior specification. 
For the slab component, we consider $\tau_1$ to be large so that $\tau_1\ge 1$. This allows the prior for $(\beta_j,\gamma_j)$ to be flat, besides the spike component at the origin.. 
We illustrate the components of a bivariate spike and slab prior in 
\cref{fig:ssbl} (with fixed $\sigma=1$). We generate the spike component 
with $\tau_0=0.001$ and the slab component with $\tau_1=5$.

\begin{figure}[h]
	\begin{center}
		\includegraphics[width = 0.95\linewidth]{spike_slab_bi.pdf}
	\end{center}
	\caption{Spike (left) and slab (right) components of a bi-variate distribution for $\tau_0 = 0.001$, $\tau_1$ = 5 and $\sigma=1$.}
	\label{fig:ssbl}
\end{figure}

For the precision term $1/\sigma^2$, a natural choice of prior is the gamma distribution
as it allows the control of both the location and the scale of the precision. To
ensure that the prior is able to represent the data, we consider $b=1$ and 
fix $a$ so that it represents the prior mean of the precision. In cases where we have no prior information,
we can simply consider a large value for $a$ so that the interval
$[0, 2a]$ contains the true value of the precision.
As defined earlier, $\pi_j$ is used as the selection probability
of the $j$-th predictor in either of the models and we use a beta prior to specify 
these selection probabilities where  $q_j$ represents our prior expectation of the selection probability ($\pi_j$) and $s$ acts as 
a concentration parameter.
For the causal effect, we want to use a Gaussian distribution that 
matches the scale of the noise term. Therefore, we consider $\beta_T\sim \normal(0,\sigma^2)$.

In \cref{fig:regress}, we show a probabilistic graphical representation
of our hierarchical model. In the figure, grey circular nodes represent the
prior hyper-parameters which will be used for sensitivity analysis
of the model. The transparent circular nodes are used to denote
the modelling parameters which are our quantities of interest. 
The observed quantities are denoted with transparent rectangular
nodes. We also use a grey rectangular node to denote the intermediate
latent variable $T^*$. We use directed edges to denote the
relationship between different nodes. However, we use a dashed
edge between $X$ and $T$ as they are related through the latent
variable $T^*$.
